---
title: "lab 11"
author: "Chan-yu Kuo"
format: pdf
editor: visual
---

Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
nrow(expr)
```

sample size for each genotype . the median expression levels for each genotypes (A/A, A/G, and G/G)

```{r}
table(expr$geno)
AA_median<-median(expr$exp[expr$geno=="A/A"])
AG_median<-median(expr$exp[expr$geno=="A/G"])
GG_median<-median(expr$exp[expr$geno=="G/G"])
AA_median
AG_median
GG_median
```

q14

```{r}
library(ggplot2)
ggplot(expr)+aes(geno,exp,fill=geno)+geom_boxplot(notch=TRUE)
```

q14: What can you infer?

The expression value on average is higher for A/A than G/G. anomalies is also more common in A/A than on G/G. Based on the median expression value, as well as the general distribution of the expression level, the SNP does effect the expression fo ORMDL3. When having an SNP (From A to G), the expression value goes down
